# SIMPLE PLOTS: BoutrosLab.plotting.general
## 2a. Create a simple scatterplot using BoutrosLab.plotting.general.
**Use the 'cars' dataset provided the R Datasets package**\
**Set a title and x- and y-axis labels that make sense**\
**Ensure that the axes ranges make sense**\
**Fiddle with other parameters as needed to make the figure aesthetically pleasing (e.g. font size, line width, etc)**\

```{r}
scatterplot.mpg.displacement <- BoutrosLab.plotting.general:::create.scatterplot(
    data = mtcars,
    formula = disp ~ mpg,
    main = 'BoutrosLab.plotting.general:::create.scatterplot mtcars',
    main.cex = 1.5,
    xlab.label = 'MPG',
    xlimits = c(8.5, 37.5),
    xat =  seq(10, 35, 5),
    xlab.cex = 1.5,
    ylab.label = 'Displacement',
    ylimits = c(50, 500),
    ylab.cex = 1.5
    );

#scatterplot.mpg.displacement;
```

## 2b. Try another function now. Create a heatmap displaying data found in the 'Loblolly' dataset found in the R Datasets package.
**Note: you may need to reorganize the data**\
**Do all formatting necessary to improve presentation (e.g. colour scheme, etc)**\

https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/Loblolly.html
The 'Loblolly' dataset contains
* height: a numeric vector of tree heights (ft).
* age: a numeric vector of tree ages (yr).
* Seed: an ordered factor indicating the seed source for the tree. The ordering is according to increasing maximum height.

```{r}
# Check Loblolly
str(Loblolly);

# Reorganize the data so that we form a matrix with age as x axis, height of each seed as y axis. Namely, we align the heights of seeds by ages.
Loblolly.reshaped <- reshape(
    Loblolly,
    idvar = 'age',
    timevar = 'Seed',
    direction = 'wide'
    );

# Check Loblolly.reshaped
str(Loblolly.reshaped);

# Use BoutrosLab.plotting.general:::create.heatmap to create heatmap
BoutrosLab.plotting.general:::create.heatmap(
    # Remove the age column
    x = Loblolly.reshaped[,-1],
    main = 'Loblolly growth heatmap by seed and age',
    main.cex = 1.5,
    xlab.label = 'Age',
    xlab.cex = 1.5,
    xaxis.cex = 0.75,
    ylab.label = 'Seed',
    ylab.cex = 1.5,
    yaxis.cex = 0.75,
    # Define the y axis labels
    yaxis.lab = (c(substring(colnames(Loblolly.reshaped[,-1]), 8, 10))),
    # Define the x axis labels
    xaxis.lab = c(row.names(Loblolly.reshaped)),
    # Don't rotate the x axis labels
    xaxis.rot = 0,
    # Show the grids
    grid.row = TRUE,
    grid.col = TRUE,
    # Don't show the clustering lines
    clustering.method = 'none',
    # Don't show the key heatmap
    print.colour.key = FALSE
    );
```

## 2c. Take a look at the 'ChickWeight' dataset in the R Datasets package
**Think about which plot-type should be used to effectively highlight the relationship between the growth of chicks and their diet.**\
**Create the plot using BoutrosLab.plotting.general functions**\
**Note that you have many options here, including plotting all the data points, aggregating the data, controlling for certain factors, etc.**\

https://rpubs.com/YaRrr/WPA0Answers
```{r}
# Get familiar with the data
str(ChickWeight);
head(ChickWeight);
```

We can create a histogram of weights for each diet
```{r}
par(mfrow = c(2, 2));

for (diet.i in 1:4) {
    hist(
        x = ChickWeight$weight[ChickWeight$Diet == diet.i],
        xlab = 'weights',
        xlim = c(0, 400),
        main = paste('Chick Weights\nDiet ', diet.i, sep = '')
        );
    }
```

A better plotting that I can think of is violin plotting.
We can use the 'yarrr' lib.
```{r}
# install.packages('yarrr');

library(yarrr);

# yarrr.guide();

pirateplot(
    formula = weight ~ Diet,
    data = ChickWeight,
    main = 'Chicken weights by Diet'
    );
```

Or use BoutrosLab.plotting.general:::create.violinplot
```{r}
BoutrosLab.plotting.general:::create.violinplot(
    formula = weight ~ Diet,
    data = ChickWeight,
    main = 'Chicken weights by Diet'
    );
```

The plots show that Diet4 yields most chicken of similar weights (around 140), which is good for standard industrial production.

