# Q2
**In bioinformatics it is often hard to identify an appropriate analytical statistical test.**
**This may be true because assumptions are violated,or because no standard test exists. In either case,a common solution is to use permutation-based approaches,like a bootstrap.**

**In this problem you will be comparing a number of different statistical approaches for comparing two groups.**
**The biological question here involves two tumour subtypes,A and B.**
**You have three samples of subtype A and nine samples of subtype B.**
**You have used a small microarray platform to measure the mRNA levels of 500 genes on each of these 12 tumours.**
**You wonder which genes differ between the two tumour subtypes.**

**Your first input data file (input1.txt) contains one column for the gene identifier and three columns of numeric data.**
**Your second input data file (input2.txt) contains one column for the gene identifier and nine columns of numeric data.**
**For both files,each column represents a biological sample (human tumour) and each row represents a gene.**
**The three numeric columns in input1.txt represent tumour-type A and the nine numeric columns in input2.txt represent tumour-type B.**
**In statistical terms,you need to determine if the three columns in input1.txt represent a random sample from the overall data for each gene.**

**Your first steps to answer this question are:**

## 1. Read the two input files
```{r}
df.subtype.A <- read.table('/Users/yupan/Downloads/new-members-documentation-training-and-guides/r-training/1. statistical-analysis/input-files/input1.txt',header = TRUE);
df.subtype.B <- read.table('/Users/yupan/Downloads/new-members-documentation-training-and-guides/r-training/1. statistical-analysis/input-files/input2.txt',header = TRUE);
```

## 2. Combine the two files into one file that contains the data for all 12 tumours.
**Make sure that the three columns in input1.txt precede the nine columns in input2.txt. Do this in the following two ways,and verify that they produce the same result:**
**a) Sort each file individually,and then use the cbind function**
**b) Use only the merge function**

```{r}
# Install dplyr,which includes a bunch of functions that make data frame operation easier.
# install.packages('dplyr',repos = 'http://cran.us.r-project.org');
library(dplyr);

# Sort both df.subtype.A and df.subtype.B by 'GeneID' in asc order.
df.subtype.A.sorted <- df.subtype.A[order(df.subtype.A$GeneID),];
df.subtype.B.sorted <- df.subtype.B[order(df.subtype.B$GeneID),];

# Combine df1.sorted and df2.sorted,remove the duplicated 'GeneID' in df2.sorted.
df.combined.by.cbind <- cbind(df.subtype.A.sorted[,1:4],df.subtype.B.sorted[,2:10]);

# Check the combined data frame
str(df.combined.by.cbind);

# The merge() function saves the troubles of sorting,joining.
df.combined.by.merge <- merge(df.subtype.A,df.subtype.B);

# all_equal() function proves that df.combined.by.merge,df.combined.by.cbind are same.
all_equal(df.combined.by.merge,df.combined.by.cbind);
```

## 3. Perform a t-test comparing the first three tumours to the last nine tumours for *each* gene using a for-loop
```{r}
# Declare the sequences instead of using magical numbers.
seq.subtype.A <- 2:4;
seq.subtype.B <- 5:13;
seq.A.B <- 2:13;

# Create a vector to store all the p_values.
res.p.values.ttest <- c();

# Get the of the rows in the merged dataframe
count.df.combined.by.merge <- nrow(df.combined.by.merge);

# Loop through each row of df.combined.by.merge
for (i in 1:count.df.combined.by.merge) {
    vector.subtype.A <- df.combined.by.merge[i,seq.subtype.A];
    vector.subtype.B <- df.combined.by.merge[i,seq.subtype.B];
    res.ttest <- t.test(vector.subtype.A,vector.subtype.B);
    res.p.values.ttest <- c(res.p.values.ttest,res.ttest$p.value);
    }
```

Using *apply functions can avoid using for-loop and make the codes more modularized and readable.

## 4. Plot a histogram of the p-values
```{r}
hist(
    res.p.values.ttest,
    main = 't-test p_values - subtype.A vs subtype.B',
    xlab = 'p-values',
    breaks = seq(0,1,0.01),
    xaxp = c(0,1,20),
    ylim = c(0,15)
    );
```

## 5. Are your axis labels rotated 90 degrees?  If so,fix this.
https://www.tenderisthebyte.com/blog/2019/04/25/rotating-axis-labels-in-r/

```{r}
hist(
    res.p.values.ttest,
    main = 't-test p_values - subtype.A vs subtype.B - y axis label vertical',
    xlab = 'p-values',
    breaks = seq(0,1,0.01),
    xaxp = c(0,1,20),
    ylim = c(0,15),
    yaxt = 'n' # this removes y-axis
    );

# redraw the y-axis
axis(
    side = 2,
    las = 2,
    mgp = c(3,0.75,0)
    );
```

## 6. Your histogram might look a bit weird in normal space,consider plotting it in log-space

The term 'log-space' here has 2 possible explanations:

1,Use the log scale,which doesn't change the frequencies but 'magnify' small scales. This can be turned on by using 'ylog = TRUE'. 

```{r}
hist(
    res.p.values.ttest,
    main = 't-test p_values - subtype.A vs subtype.B - ylog = TRUE',
    xlab = 'p-values',
    breaks = seq(0,1,0.01),
    xaxp = c(0,1,20),
    ylim = c(0,15),
    ylog = TRUE
    );
```

But in this sample,the frequencies are not big,using it doesn't make any difference. So probably the next explanation is what the exercise desires.

2,Apply the logarithmic conversion to the frequencies. This also 'magnifies' small frequencies. 
Below is how to implement it.
```{r}
# Get the original counts of all bins
myhist <- hist(
    res.p.values.ttest,
    breaks = seq(0,1,0.01),
    plot = FALSE
    );

# Apply logarithmic to the counts
myhist$counts <- log(myhist$counts,10);

plot(
    myhist,
    main = 'Ttest p-values with log10(frequency)',
    xlab = 'p-values',
    xaxp = c(0,1,20),
    ylab = 'log10(Frequency)',
    ylim = c(0,1.5),
    col = 'grey'
    );
```

## 7. What does this distribution tell you?
The p_values are relatively evenly distributed in the range (0,1).
It'd be interesting to look into those genes that gave p_values < 0.1,also check the effect sizes,in order to find out the genes that really differentiate input1 and input2.

