# Q6
**To this point all the analyses have been performed using standard R functions.**  
**Fortunately the BoutrosLab repository has a variety of functions that will aid in making the plots and statistics.**  
**Modify your code to use:**
**a) BoutrosLab.statistics.general for all p-value extraction functions**
**b) BoutrosLab.plotting.general for all plots**

```{r}
# Call get_merged_df.R first to get the required data
source('get_merged_df.R');

# Overwrite the seq.subtype.A and seq.subtype.B defined in get_merged_df.R
seq.subtype.A <- 1:3;
seq.subtype.B <- 4:12;
```

## a) BoutrosLab.statistics.general for all p-value extraction functions
```{r}
# Use BoutrosLab.statistics.general::get.ttest.p
res.p.values.ttest.by.BL <- apply(
    X = df.combined.by.merge[,seq.A.B],
    MARGIN = 1,
    FUN = function(x) {
        BoutrosLab.statistics.general::get.ttest.p(
            x,
            group1 = c(seq.subtype.A),
            group2 = c(seq.subtype.B)
            )
        }
    );

# Use BoutrosLab.statistics.general::get.utest.p
res.p.values.utest.by.BL <- apply(
    X = df.combined.by.merge[,seq.A.B],
    MARGIN = 1,
    FUN = function(x) {
        BoutrosLab.statistics.general::get.utest.p(
            x,
            group1 = c(seq.subtype.A),
            group2 = c(seq.subtype.B)
            )
        }
    );

# Use BoutrosLab.statistics.general::get.foldchange
fold.changes.by.BL <- apply(
    X = df.combined.by.merge[,seq.A.B],
    MARGIN = 1,
    FUN = function(x) {
        BoutrosLab.statistics.general::get.foldchange(
            as.numeric(x),
            group1 = c(seq.subtype.A),
            group2 = c(seq.subtype.B)
            )
        }
    );
```

## b) BoutrosLab.plotting.general for all plots
```{r}
# Use BoutrosLab.plotting.general::create.histogram to draw histogram for res.p.values.ttest.by.BL
BoutrosLab.plotting.general::create.histogram(
    x = res.p.values.ttest.by.BL,
    main = 'ttest p-values histogram by BoutrosLab.plotting.general::create.histogram',
    main.cex = 1,
    xlab.label = 'p-values',
    xlab.cex = 1,
    type = 'count',
    xlimits = c(0,1),
    ylimits = c(0,15),
    breaks = seq(0,1,0.01),
    xat =  c(seq(0,1,0.1))
    );

# Use BoutrosLab.plotting.general::create.histogram to draw histogram for res.p.values.utest.by.BL
BoutrosLab.plotting.general::create.histogram(
    x = res.p.values.utest.by.BL,
    main = 'utest p-values histogram by BoutrosLab.plotting.general::create.histogram',
    main.cex = 1,
    xlab.label = 'p-values',
    xlab.cex = 1,
    type = 'count',
    xlimits = c(0,1),
    ylimits = c(0,80),
    breaks = seq(0,1,0.025),
    xat =  c(seq(0,1,0.1))
    );

# Use BoutrosLab.plotting.general::create.histogram to draw histogram for fold.changes.by.BL
BoutrosLab.plotting.general::create.histogram(
    x = fold.changes.by.BL,
    main = 'fold changes histogram by BoutrosLab.plotting.general::create.histogram',
    main.cex = 1,
    xlab.label = 'fold changes',
    xlab.cex = 1,
    type = 'count',
    breaks = 100
    );
```


